---
title: 'Allies as Armaments: Explaining the Specialization of State Military Capabilities'
author:
  - name: Anonymized author
format:
  pdf:
    fig-format: png
    thanks: Acknowledgements omitted for blinded peer review.
    number-sections: true
    fontsize: "12"
editor: visual
execute:
  echo: false
  warning: false
  cache: true
date: today
date-format: long
abstract: \singlespacing \noindent Scholars and practitioners have long maintained that, for capable states, a full-spectrum military provides the best security in an unpredictable anarchic world. However, not all states seem to make this choice, with some capable states have imbalanced, specialized militaries. Are specialized militaries that forgo the development of vital defense capabilities and overinvest in seemingly less necessary ones simply making mistakes? I argue that there are conditions under which states can reap the gains of economic efficiency that stem from specialization without sacrificing the security benefits of a full-spectrum force. Because defense alliances build trust, minimizes the risk of defection, and foster coordination, states with militarily-capable alliance partners can individually specialize in military capabilities that, when brought together, still comprise a full-spectrum military force. I substantiate these arguments with a new measure of military portfolio specialization using fine-grained data on state military assets from 1970-2014.
header-includes:
- \usepackage{tikz}
- \usepackage{pgfplots}
- \pgfplotsset{compat=newest}
- \usetikzlibrary{plotmarks}
- \usetikzlibrary{arrows.meta}
- \usepgfplotslibrary{patchplots}
- \usepackage{grffile}
- \usepackage{caption}
- \usepackage[utf8]{inputenc}
- \usepackage[doublespacing]{setspace}
- \AtBeginEnvironment{tabular}{\singlespacing}
- \usepackage{float}
- \usepackage{multirow}
- \usepackage{tablefootnote}
- \usepackage{pifont}
- \usepackage{newunicodechar}
- \usepackage{booktabs}
- \usepackage{siunitx}
- \usepackage{tabularx}
- \newunicodechar{✓}{\ding{51}}
bibliography: ForceStructure.bib
---

```{r knitr_options}
library(knitr)
library(kableExtra)
library(magrittr)
library(ggplot2)
library(ggtext)
```

```{r load data}
df_raw <- readRDS(paste0(here::here(), '/data/rDMC_raw_v1.rds'))

df_long <- readRDS(paste0(here::here(), '/data/rDMC_long_v1.rds')) %>%
  dplyr::filter(year <= 2014)

df_wide <- readRDS(paste0(here::here(), '/data/rDMC_wide_v1.rds')) %>%
  dplyr::filter(year <= 2014)

df_full <- readRDS(paste0(here::here(), '/data/03_DF-full.rds'))
```

# Introduction {#sec-intro}

Despite constitutional restrictions on its military capabilities, Japan began shifting its military investments in the late 1970's. By 1982, Prime Minster Suzuki had draw up plans to overhaul Japan's military by investing primarily in air defense and light offshore surface ships [@modly_rhetoricrealitiesjapan_1985]. Although the Soviet threat in East Asia ended a decade later, by the turn of the century Japan had doubled its air defense and short-range aerial capabilities and almost completely phased out its amphibious fleet and coastal ships, despite their clear utility for an island-chain state. Japan's military portfolio had changed fairly significantly by specializing in some capabilities and omitting others in a way that did not seem to align with the international security threats they were facing.

This is exemplary of a broader phenomenon of interest to international relations scholars. Why do some states possess seemingly sub-optimal militaries - under-producing some capabilities or over-producing others - while other states maintain a more diversified military portfolio? How states plan for conflict and the military tools this involves have much to tell us about war's causes and consequences. The combination of capabilities that comprise a military's toolkit determine the operations it plans for and undertakes, the types of threats it can credibly make, and the consequences of resorting to force [@buzan_introductionstrategicstudies_1987]. Assuming that the primary purpose of a state's military is to provide security against perceived threats [@waltz_theoryinternationalpolitics_1979, 102-114], most states should optimize their force structure by diversifying their capabilities because it compensates for the inherent weaknesses of any one set of capabilities and helps deal with unanticipated or opaque threats [@biddle_militarypowerexplaining_2005, 199-200]. And yet, there are myriad examples of seemingly risky specialization like the US forgoing minesweepers in the 1980's even though 13 of the 15 US ships sunk since World War II were victims of naval mines [@till_holdingbridgetroubled_2005], Albania in the early 2000's producing dozens of patrol vessels that could be deployed as far as western Portugal [@polak_natomembershipalbania_2009], or Estonia investing in sophisticated cyber capabilities but having no combat aircraft despite warranted concern about Russian aggression [@andzans_deterrencedilemmalatvia_2017]. Why do some capable countries have gaps in their militaries that they could fill, but choose not to?

My central argument is that variation in the specialization or diversification of a state's defense capabilities can, in part, be explained by the presence of militarily-capable allied states. Alliance relationships allows states to reduce the cost of forgoing some capabilities and overproducing others. States can have their cake and eat it, garner the benefits of having both specialized and diversified military capabilities by individually specializing when the relationship with their allies can produce collective diversification. The result is variation in the composition of military capabilities across states - some being comparatively more specialized or diversified. I substantiate these arguments by examining the relationship between a state's military specialization and the nature of its alliance relationships from 1970 - 2014, and find that states with more militarily-capable alliance partners are able to specialize their militaries more than those with weak or non-existent defense partnerships.

In the next section, I detail existing research concerning the factors that determine a state's force structure in general, and more specifically why capable states sometimes pursue a specialized distribution of military capabilities, despite the security risks that engenders. @sec-theory then introduces a shared production model of defense, theorizing that alliances can sufficiently minimize the risks inherent in military specialization. @sec-empirics provides an empirical test of this theory using a new entropy-based measure of a state's military portfolio specialization adapted from statistical ecology and applied to existing data on disaggregated national military capabilities. @sec-conclusion concludes by discussing the implications of these findings and motivating future research on the characteristics of alliances that explain why they produce divisions of labor that vary in degree and in kind.

# Existing Explanations for Variation in the Distribution of Military Capabilities {#sec-lit}

Most recent research on military capabilities focuses on variation in the *size* of state militaries [@cappellazielinski_whatgoesmust_2017; @blankenship_controlvscostsharing_2018]. However, such research does not explain why militaries vary in their *composition*. Early discussions about the political determinants of a state's weapons development were framed around a debate between internal and external causes [@evangelista_innovationarmsrace_1988]. Theorists forwarding internal explanations argued that because there was no single authority for weapons development decisions [@allison_armamentsarmscontrol_1975; @kossiakoff_conceptionnewdefense_1980], the composition of a state's military was determined by domestic factors like bureaucracy [@farrell_weaponscausepolitics_1997], constituency interests [@higgs_hardcoalsmake_1988], or scientific R&D culture [@zuckerman_nuclearillusionreality_1982]. In contrast, the external causes theorists argued that weapons acquisitions decisions were primarily motivated by a strategic response to foreign threats [@mcnamara_remarkssecretarydefense_1967].

## Internal Sources of Armament Decisions

Theories of internal sources of armament decisions have typically tried to explain weapons acquisition more generally, rather than identifying whether those weapons acquisitions are consistent with a specialized or diversified aggregate military portfolio. These theories examine factors like the economic motivation to support defense contractors [@kurth_whywebuy_1973], thus identifying the importance of the military industrial complex in shaping defense policy [@kotz_wildblueyonder_1988], although others believe this is limited by a regulatory environment that conditions lobbying efforts [@mawdsley_armamentsdecisionmakingare_2018]. Separately, re-election incentives may explain weapon developments that generate jobs or shore up nationalism [@smith_powergamehow_1988; @carter_senatedefensebudgeting_1989] although there is disagreement about the empirical record [@lindsay_testingparochialhypothesis_1991]. Similarly, political ideology and regime type may shape preferences for or against a particular military capability, as evidenced by trade protectionist support for battleship fleet development [@kehr_battleshipbuildingparty_1975; @fordham_domesticpoliticsworld_2019] and autocratic aversion to weapons that could be used in a coup [@brown_recoupingcoupproofingcompromised_2016].

More socially-driven domestic considerations point to the importance of non-state actors and incentives, but are less tied to the assumption of egoistic profit motivations and political self-interest. Instead, the weapons a state develops may be decided by scientists and technologists [@zuckerman_nuclearillusionreality_1982], although this perspective has been challenged by further empirical examinations of the same Cold War case studies [@mackenzie_inventingaccuracyhistorical_1993; @spinardi_polaristridentdevelopment_1994]. More sociological theories have posited that status concern explains particular weapons acquisitions like high-technology aircraft or naval carriers [@eyre_statusnormsproliferation_1996; @hintz_symbolicamplificationsuboptimal_2022] but only in limited empirical cases.

## External Sources of Armament Decisions

The dominant "external cause" explanation for why a state has the military capabilities it does concerns that state's international threat environment [@nordhaus_effectsinternationalsecurity_2012], meaning variation in states' distributions of military capabilities is explained by the perceived best response to security threats, subject to economic capacity [@resende-santos_neorealismstatesmodern_2007]. One important implication of this neo-realist assumption about the self-help international system concerns states as like-units [@parent_balancingneorealism_2015]. States with similar resource endowments should have similar distributions of military capabilities even when they face a common enemy [@resende-santos_anarchyemulationmilitary_1996; @goldman_systemiceffectsmilitary_1999]. @waltz_theoryinternationalpolitics_1979 [p. 127] argues that even where material capabilities differ, "contending states imitate the military innovations contrived by the country of greatest capability and ingenuity. And so the weapons of major contenders, and even their strategies, begin to look much the same all over the world." States facing similar threats should have similar militaries because "contemplation of a common fate may not lead to a fair division of labor - or to any division of labor at all" [@waltz_theoryinternationalpolitics_1979, 165]. Because states cannot resolve the problem of credibly relying on one another, the self-help nature of the international system should prevent states from being able to functionally differentiate their military capabilities by relying on each other. While there is "functional specialization - an intense division of labor" *within* states and other organizations, "\[t\]he same is obviously not true of international politics. There, power is distributed more equally than in organizations. Moreover, it is distributed to protect no group purpose. There is no functional specialization among states." [@posen_sourcemilitarydoctrine_1984, 36-37].

Since the absence of an international sovereign makes cooperation under anarchy difficult, states try to maximize their security through a full-spectrum approach to defense where each states individually acquires the military capabilities they deem necessary (and feasible) for their national security. This optimal distribution of capabilities is assumed to be a diversified one because "each weapon, unit, and technique possesses a unique set of capabilities and vulnerabilities. Taking full advantage of these military assets increases the likelihood that an armed force will fulfill its mission" [@millett_effectivenessmilitaryorganizations_1986, 52]. The historical record also finds this logic among practitioners, as "a synergistic mix of platforms and weapons intended to produce a force or more capability than that represented merely by the sum of its parts, has always been justified by the argument that it provides a wide range of responsive options" [@till_maritimestrategytwenty_1994].

## Unexplained Variation in the Aggregate Composition of Military Capabilities

While domestic politics influences many aspects of force structure like acquisition decisions, production capacity, and innovation patterns, the predictions they generate for the overall composition of a state's military are less clear. These theories do not try to explain why, for example, states with an influential military-industrial complex would have highly specialized force structures or why we should expect states with divided governments or protectionist politicians to have a more or less diversified force. Similarly, while threats certainly condition what capabilities a state chooses to arm itself with, these theories do not explain the empirically observed decision by capable states to specialize their militaries by omitting or over-producing capabilities given their threat environment. They assume states are like-units that, holding economic capacity and geography constant, should converge in developing similarly diversified full-spectrum militaries. Sameness and diversification are theoretically assumed, but not empirically substantiated.

# A Theory of Specialization Within Alliances {#sec-theory}

States face a constrained optimization problem where the set of resources available to accomplish a given task are finite. So when allocating resources to defense, decisions about the capabilities a state should possess involve a zero-sum balance between efficacy and efficiency. States can manage this trade-off by allocating resources toward many capabilities (prioritizing efficacy) or toward a few (prioritizing efficiency). States prioritizing efficacy invest in a diversified military portfolio because it reduces their overall vulnerability, but at a relatively higher economic cost. In contrast, a state prioritizing efficiency invests in a specialized military portfolio because it comes with economies of scale and improved integration, but at the risk of not having capabilities it may need.

Business organization research on inter-firm cooperation has identified a way to manage this trade-off [@gulati_strategicnetworks_2000]. Strategic alliances with like-minded actors can minimize the otherwise zero-sum trade-off between efficacy and efficiency by allowing functional specialization via a division of labor. When this occurs, states can overcome part of the constrained optimization problem by individually specializing (efficiency gains) while collectively diversifying (efficacy gains). Specialization by capable states is thus not a questionable prioritization of efficiency at the expense of efficacy by states choosing to forgo the security benefits of a diversified military, but instead an optimality made possible by architectures of international cooperation.

This theory is motivated by a common set of assumptions about rational social choices by states.[^1] States are rational insofar as their preferences are transitive, their actions purposive, and their criteria for evaluating interstate relations is instrumental rather than innate [@frieden_actorspreferencesinternational_1999]. In assuming that states act rationally, I do not presuppose complete information. States still suffer unintended consequences and make mistakes given uncertainty about the state of the world, including the true underlying values of the variables of interest that they regard with some unknown probability distribution. While recognizing that state preferences "arise in the first place from bargaining solutions to problems of order" [@mattern_hierarchiesworldpolitics_2016, 634], I remain agnostic about the origins of these preferences as I am not trying to explain where state preferences come from.[^2]

[^1]: For contrasting views on the benefits and drawbacks of assuming rational decision-making in the specific context of the economics of arms acquisition, see @sandler_politicaleconomynato_1999 and @fevolden_defenceindustrialpolicy_2016.

[^2]: For parsimony, I largely bracket the decisions and preferences of domestic actors and instead treat states as units with homogeneous preferences. Domestic considerations clearly play a role in state defense policy as an input into both the dependent and independent variable, but not in a way that invalidates the explanatory role of interstate relationships.

## Costs and Benefits of Specialization

To justify explaining variation in the distribution of capabilities, I first identify the costs and benefits of both ends of the dimension of interest - specialization and diversification - summarized in @fig-spectrum_specialization. Although the benefits of military specialization initially seem like economic issues that should take a backseat to security considerations, the two are inevitably intertwined because a state's decisions about how to best provide for its defense occur within a constrained optimization environment. Thus, economically-conscious defense decisions impact how well a state will be able to provide for its security and how well various aspects of their defense portfolio work with one another during conflict. The three primary benefits of specialization stem from economies of scale, operational efficiency, and improved integration [@meier_culturetrustdivision_2019].

```{r, engine='tikz', echo=FALSE}
#| label: fig-spectrum_specialization
#| fig-cap: "Varieties of a state's distribution of military capabilities."

\begin{tikzpicture}

  % Baseline
    \draw [<->, thick] (0,0) -- (8,0);
                    
    % Labels            
    \draw (0,0) node[above = 3pt, align = center] {Diversification} 
        node[below = 3pt, align = center] {$\cdot$ \small Available capabilities \\ $\cdot$ \small Reduced vulnerability};
    \draw (8,0) node[above = 3pt, align = center] {Specialization} 
        node[below = 3pt, align = center] {$\cdot$ \small Economies of scale \\ $\cdot$ \small Operational efficiency \\ $\cdot$ \small Improved integration};
                    
    % Headers
    \draw (-4,0) node[above = 3pt, align = center] {\textbf{Distribution of} \\ \textbf{Military Capabilities}} 
        node[below=9pt, align = center] {\textbf{Benefits}};

\end{tikzpicture}
```

First, the cost of setting up manufacturing and industrial plants as well as acquiring the materials for weapons acquisition often entail large up front investment. But the marginal cost of that investment goes down as a state decides to produce more of the same asset [@markowski_challengesdefenceprocurement_1998]. For example, Germany has reduced the need for redundant infrastructure by centralizing car and light truck production all within the Bundeswehr-Fuhrparkservice GmbH which allows them to produce newer vehicles more quickly, although of less variety [@overhage_poolitshare_2013]. This is still true for states that are primarily "arms buyers" as opposed to "arms builders", as maintenance and repair costs are reduced when there is a smaller list of components and end-use products involved. Economies of scales are also"active" in that they accrue as a state undertakes defense-related activities, so the more a state operates with a particular asset, the lower their marginal costs because of "learning by doing" [@postrel_islandssharedknowledge_2002].

Second, specialization allows a military to perform select missions more efficiently by streamlining logistics and reducing the overall cost of learning how to use new equipment. Many assets require capability-specific investments that involve a fixed cost. A state with several dozen different types of aircraft will require more complex pilot training than a state that only has to master the effective use of a few types of aircraft. One source of the internal debate over which NATO members should send main battle tanks to Ukraine concerns the Ukrainian military's familiarity with how those more complex systems work; they could immediately operate T-72 tanks sent from Eastern Europe, but training and logistics for the US Abrams tank would take months [@lanoszka_artpartialcommitment_2023, 6-10].

Third, integration is easier as a country specializes since the complexity of integrating numerous types of platforms with various roles and responsibilities decreases. Even issues as fine-grained as the software used in various pieces of equipment are sufficient impediments to military operations that nations consider this issue carefully. NATO's Standardization Agreement (STANAG), for example, ensure broad fleet compatibility with the same fuel nozzle. Lest this seem like a story about NATO, in 2019 Jordan gave up its Chinese-built CH-4 drone fleet because successful integration with other platforms was going to require a costly overhaul of their communication systems [@penney_modernizinguavexport_2020].

In contrast to specialization, the benefits of diversification manifest as the security gains of a full-spectrum military that makes combined arms warfare possible [@biddle_militarypowerexplaining_2005]. States that engage in a full-spectrum approach to warfare instead of specializing benefit from having all the capabilities needed to defend themselves. No weapons system is perfect, and the nature of warfare means weapons systems that excel at one aspect of international conflict do so precisely because they lack other capabilities. Aerial bombers can sacrificed maneuverability so that they have carry a high payload. More maneuverable aircraft like fighters have to compensate for the benefits of speed and maneuverability with lower ordinate payloads in order to achieve those benefits.

Diversification also reduces vulnerability because failure to diversify makes it easier for the adversary to develop countermeasures. A state with a limited variety of military capabilities has given their adversaries a shorter list of defense capabilities they must be able to defeat to prevail in conflict. Air defense systems, for example, come in three different varieties; surface-to-air missiles (SAM), anti-aircraft artillery (AAA), and aircraft armed with air-to-air missiles (AAM). These systems all differ in the altitudes they can target, stealth, reaction times, mobility, and cost. In a 1940 testimony before the Senate Appropriations Committee, General George Marshall noted the need for both aircraft and anti-aircraft artillery because the former is an area system that excels at searching while the latter is a point system designed to protect key assets. When asked by Congress which was preferable, he said "the whole thing is interwoven...all these matters have to be given proper weight to get a well integrated and balanced whole" [@hammel_roadbigweek_2010]. A state that has chosen to develop only one of these capabilities might have more in quantity (scale economies) and quality (operational efficiency and improved integration), but they are now vulnerable to the development of new missiles and aircraft designed to circumvent the strengths of their adversary's one air defense system.

States that have not embraced diversified militaries have consequently suffered. After the Yom Kippur war, Israel opted to specialize their military by cutting artillery and mechanized infantry in favor of a shift to pure armor-aircraft. This left them vulnerable to an anti-armor and anti-aircraft attack that set them back in the early stages of the 1973 war against Egypt. It was only after they reversed course that they were able to defeat the Egyptian air defense systems [@herzog_waratonementstory_2018]. Operating with a similar logic, India's primary defense planning document after World War II, the @navalplanpaper_1947, also made the case for a "balanced naval task force" which was later clarified by Indian Vice Admiral W.E. @parry_indiaseapower_1949 as "containing all types of ships and aircrafts, on the sea, over the sea, and under the sea" in a move to reduce India's vulnerability.

## Individual Specialization Within Alliances

Given the relative costs of a diversified or specialized distribution of military capabilities, it appears the wise decision is to diversify as best as possible given an environment constrained by limited material resources. So, why do states specialize? In particular, why do some states consciously choose to forgo optimal defense simply to capture economies of scale? Doing so appears to be a decision to choose a sub-optimal defense posture that could be have been avoided. Specializing one's military is not sub-optimal if it is part of a cooperative division of labor that means a state gets access to a diverse range of military capabilities provided by allied states. After all, one way states can save resources in a constrained optimization environment is by sharing the production of international security with other states in a way that allows each state to allocate resources toward non-security functions [@digiuseppe_armsdemocraticallies_2016]. This can garner the security benefits of capability aggregation posited by the neo-realists as well as the economic benefits put forth by hierarchy theorists.[^3]

[^3]: @auerswald_natoafghanistanfighting_2014 [p. 229-233] also identify how specialization by alliance members can help individual states engage in cost-effective defense investments while maintaining an aggregate full spectrum combined-arms force, but reach the opposite conclusion in remaining skeptical that the risk of opportunism and cost of coordination can be adequately addressed. Their analysis is driven by European debates over the Smart Defence initiative that stalled due to fears that participating countries could refuse to participate in coalition warfare.

A state's decision about whether to cooperate with other states over the ownership and use of security assets is a function of three factors - the gains from cooperation, the expected cost of opportunism, and the expected cost of coordination [@lake_risefallfuture_1997]. When the gains of cooperation exceed the expected costs of opportunism and coordination, they can specialize their militaries because cooperation means doing so does not sacrifice the benefits of a interstate diversified military force. Otherwise, problems of opportunism [@dekker_controlinterorganizationalrelationships_2004] and coordination costs [@oxley_appropriabilityhazardsgovernance_1997] inhibit otherwise fruitful defense cooperation. The gains from cooperation include those previously outlined like economic benefits from economies of scale, political benefits from the efficiency gains of focusing on core competencies, and security benefits from improved performance at particular security needs. The expected cost of opportunism is a function of the severity and likelihood of abandonment, entrapment, and exploitation [@snyder_securitydilemmaalliance_1984]. The expected cost of coordination concerns how much work is required to ensure the relationship achieves the expected benefits [@gulati_architecturecooperationmanaging_1998]. When a state is confident they can rely on their allies for some defense capabilities and is willing to contribute their own capabilities to the defense of their allies, a state will specialize.

There are *gains to cooperation* in a security arrangement simply by virtue of having access to more resources. After all, one of the expected payoffs from developing relationships with similarly aligned states is the expectation that some aspect of your ally's military resources are available during war [@conybeare_portfoliodiversificationmodel_1992]. Ideally, the resource gains under cooperation would be more than the sum of its parts because there are scale economies in the production of defense through factors like standardization, repetition, fixed initial investment, and learning. Discussions in the US about a '1,000 ship Navy' are predicated on precisely this model; a navy that is "not a thousand gray hulls flying the American flag, but rather a voluntarily global maritime network that ties together the collective capabilities of free nations to establish and maintain a dramatically increased level of international security in the maritime domain" [@morgan_000shipnavyglobal_2005]. Practitioners have recognized how having allies can foster focused specialization. US Naval Rear Admiral @smith_strategiccooperationeverybody_2013 noted that by having this cooperative approach, "each nation can avoid duplication and thereby reduce its proportional share of the expense. This is not simply about global maritime partnerships. It's about a focused and pragmatic approach to force allocation that acknowledges allies' existing contributions. Countries could immediately apply the freed resources to unique national missions for which a collaborative approach is impractical."

Concerning the *risk of opportunism*, both states must feel that mutual cooperation is preferred to mutual defection in cases where unilateral defection is preferred to unrequited cooperation [@axelrod_achievingcooperationanarchy_1985; @oye_explainingcooperationanarchy_1985]. Otherwise, relying on another state that may renege when asked to contribute to your defense could seriously jeopardize a state's security. If this risk is seen as high, states should instead opt to produce security on their own. Opportunism via abandonment (shirking or buck-passing), entrapment (chain ganging), or exploitation can prove fatal to a state that depends on another for defense [@snyder_securitydilemmaalliance_1984], especially if that dependence took the form of a specialization via omitting necessary defense assets. Avoiding this requires avenues for communication and routines for interaction that mitigate concerns about opportunism sufficient to encourage coordinated military strategies [@ikenberry_victoryinstitutionsstrategic_2001]. Because relying on your partner to provide assets risks losing autonomy over the conduct of those assets during combat, there can be ambiguity about the effects of cooperation on a state's security goals [@morrow_alliancesasymmetryalternative_1991]. As @auerswald_natoafghanistanfighting_2014 [p. 232] put it, "in alliance warfare, allies sometimes do not always show up when needed or they show up but are not able to do what is needed."

The *costs of coordination* are distinct from those incurred from opportunism [@white_cooperationcostsgovernance_2005, 1385]. While opportunism costs focus on uncertainty regarding your partner, coordination costs are about uncertainty regarding the task. You can have full confidence your partner will not act opportunistically if interests are perfectly aligned, but there still has to be coordination about who is specializing in what and a formal designation of roles. Despite this, international relations research on the problems of cooperation inspired by @williamson_economicinstitutionscapitalism_1985 and the accompanying transaction cost framework have been less concerned with coordination costs, instead arguing that opportunism costs should be at the forefront [@lake_anarchyhierarchyvariety_1996]. Coordination costs may not be particularly salient in contexts where resources are simply pooled, but that is not the case in the security context [@overhage_poolitshare_2013]. It is quite difficult to fight a war with another state's tools. In defense, "duplication of facilities, differences in requirements, coordination problems, lack of clear control and delays due to different budgetary systems all tend to increase the costs of collaborative projects" [@smith_weaponsprocurementdefence_1996, 69-70]. Collaboration requires communication, making adjustment in response to your partners actions. This is fundamentally an issue of information asymmetry, so creating a "common knowledge assumption" can induce and stabilize cooperation by reducing uncertainty about the other actor's payoffs structure and conveying one's own payoff structure to your partner [@gulati_mutuallyassuredalliances_1992, 19].

In sum, seemingly sub-optimal force structures - those that omit useful defense capabilities and/or overproduce others - can occur when a state has opted to specialize its military portfolio. A state is more willing to do so when the security risks of specialization are no longer prohibitive; a condition that is made possible by alliance relationships that reduce those security risks. Under these conditions, alliances represent a means to deal with the constrained optimization problem by taking advantage of capability specialization made possible by shared production. Specialization is thus an outcome of security alliances because they promote a division of responsibility and function.

# Empirical Analysis {#sec-empirics}

## Dependent Variable

The dependent variable is the degree of specialization of a state's distribution of military capabilities in a given year. A state's distribution of military capabilities is defined here as the combination of military equipment that could be used by a state during conflict. This includes platforms like artillery, aircraft, naval vessels, armored vehicles, satellites, and transport ships.[^4] I choose these scope conditions because military platforms are equipment that can be deployed, that other nations are likely to observe, that could be used to signal intent and resolve in a crisis without actual use, and that are durable goods. The index is constructed using the rDMC dataset, which details national military portfolios by providing annual counts of `r scales::comma(dplyr::n_distinct(df_long$tek))` different military platforms across all states from 1970-2014 [@gannon_planestrainsarmored_2021].

[^4]: The data do not include munitions like single-use bombs or ammunition or firearms used by individual military personnel. Existing research has made similar distinctions in what military capabilities are examined cross-nationally [@brooks_risefallgreat_2016; @caverley_hegemoncurveproduction_2021].

To measure military specialization at the country-year level, I create an index quantifying the differences across states' distribution of military capabilities identified as omissions and over-productions relative to the neorealist baseline assumption that states behave as like-units under anarchy and should consequently seek similarly diverse military capabilities subject to resource constraints. Assume that global defense in year $t$ is composed of $N$ countries and $M$ military technologies. I construct an $n \times m$ interaction matrix for each year $t$ such that each row $n$ is a country and each column $m$ is a technology. Each cell thus represents the observed count of a given technology in that country-year's military. In aggregate, this can be represented as $d_j = \sum_{i=1}^{N}(p'_{ij}ln\frac{p'_{ij}}{q_i})$ where $N$ is the total number of countries in that year, $p_{ij}$ is country $i$'s possession of technology $j$ divided by the total amount of technologies $j$, and $q_i$ is the total number of technologies possessed by country $i$ divided by the total number of technologies in the world.[^5]

[^5]: This bipartite network structure is modeled after its use in ecological research [@alarcon_yeartoyearvariationtopology_2008].

From this, I calculate the functional entropy of each country's military using a trait-based similarity measure based on @rao_diversitydissimilaritycoefficients_1982's quadratic entropy calculation of the average difference across technology portfolios between each country and all other countries in a given year weighting the technologies by their relative abundance [@pavoine_phylogeneticfunctionaloriginality_2017]. This calculates the functional entropy of a country's military as: $R(p_i,D) = \sum_{k=1}^{S} \sum_{l=1}^{S} \sqrt{(p_k|i)} \sqrt{(p_k|j)} d_{kl}$ where $p_i=(p_1|i, …, p_k|i, …, p_S|i)$ is the vector of relative technology abundance within country $i$; $S$ is the number of technologies; $D=(d_kl)$ is the matrix of functional dissimilarity between the technologies, and $d_kl$ is the functional dissimilarity between countries $k$ and $l$ [@pavoine_adivpackageanalyse_2020].[^6]

[^6]: As this measure of functional entropy is developed by @pavoine_phylogeneticfunctionaloriginality_2017, its formula is provided verbatim. In the original statistical ecology application, this measure uses @hill_diversityevennessunifying_1973 numbers to measure the dissimilarity between biological species based on observed traits, accounting for the rarity of those traits.

To provide some intuition, this measure of entropy calculates the degree of surprise or unpredictability produced by the difference between the amount of a military capability we expect a country to possess and what that country actually possess. This prior expectation is based on the distribution of technologies across all other states and within the state in question. For example, if most states possess, on average, twice as many transport helicopters as they do transport aircraft, we would expect a state with 10 transport aircraft to have roughly 20 transport helicopters. But if the state in question already possessed many more transport aircraft than everyone else, we would update our expectation since we know a way this quantity differs from other states and other capabilities. Our expectation for transport helicopters can thus be *re*-calibrated based on (1) the number of transport aircraft this states possesses relative to everyone else's transport aircraft, and (2) the number of transport aircraft this state possesses relative its other capabilities. If we now reproduce this method across all other capabilities, we get a revised prior expectation for the capability in question - transport helicopters. The closer the observed quantity is to our final re-calibrated expectation, the less entropy the quantity produces, and thus the lower the level of specialization since producing many more or far fewer transport helicopters than the model expects are both indications that the state has specialized by omitting or over-producing that capability relative to expectation.

@fig-spec_scores shows the distribution of this index across all observations.[^7] Among the most specialized observations are the United States in the early 2000's and Japan after 2010. The least specialized states include the Baltic states in the early 1990's and Barbados. This may initially seem counter-intuitive. After all, the conventional wisdom holds that superpowers possess full-spectrum military forces [@dombrowski_buyingmilitarytransformation_2006, 8]. Yet practitioners think about a full-spectrum force as involving the ability to "prevail, quickly, and cheaply, in any and all forms of conflict" [@powell_myamericanjourney_2010, 157]. This is not mutually exclusive with a form of specialization that emphasizes some capabilities over others. Capable states with full-spectrum militaries are still subject to constrained optimization and are unable to excel at all forms of conflict simultaneously [@lake_technologyqualitativesuperiority_2012, 91]. Given the number of submarines, frigates, and destroyers the US possesses, having roughly a dozen aircraft carriers but less than three minesweepers is an unexpected form of naval specialization. Making priorities is both a product of luxury and of necessity. So while the United States may dominate in many military capabilities, that does not mean its relative dominance is equal across the board. Similarly, having a diversified military is not synonymous with having a lot of everything. States can have very little of everything, making them similarly *in*capable across the board.[^8]

[^7]: There is a small, statistically significant, positive temporal trend that is accounted for in the statistical models.

[^8]: Nonetheless, these descriptive statistics identify the importance of the statistical model accounting for economic capacity to ensure the specialization index is not simply measuring the economic luxury of choice.

```{r}
#| label: fig-spec_scores
#| fig-cap: "Distribution of the dependent variable (country-year military specialization). The measure is bounded between 0 and 1, with higher values representing higher specialization. 1 represents the highest theoretically possible level of specialization."
#| fig-pos: 'h'

df_full %>%
  ggplot(aes(x = spec_fentrop)) +
  geom_histogram(bins = 30, color = "black", fill = "gray") +
  geom_vline(aes(xintercept = mean(spec_fentrop, na.rm = TRUE)), 
             linetype = "dashed", linewidth = 0.6) +
  annotate("text",
           x = mean(df_full$spec_fentrop, na.rm = TRUE) + 0.035,
           y = 625,
           label = paste("Mean =", round(mean(df_full$spec_fentrop, na.rm = TRUE), 3))) +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  labs(title = "Specialization of Distribution of Military Capabilities (DMC)",
       x = "Specialization",
       y = "Number of country-years") +
  theme_classic() +
  theme(text = element_text(size = 10))
```

## Independent Variable and Controls

The independent variable measures a state's alliance relationships. As most states have at least one formal treaty ally in a given year, existing research has modeled alliances as an independent variable by proxying for the importance of a state's allies to that state's security. I operationalize alliances at the country level two different ways; (1) as the logged sum of military spending of a state's allies (excluding itself) [@digiuseppe_armsdemocraticallies_2016] and (2) the ratio between a state's CINC score and the sum of their alliance CINC scores (including itself) [@fang_concederesistrestraining_2014; @johnson_capabilitycredibilityextended_2015]. For both variables, higher values indicate more important alliance relationships. These variables are both observable indicators of conditions conducive to military specialization because a state with more militarily capable allies should be more confident that specializing its military will not leave it vulnerable to threats because its allies will help provide for its defense. A state's allies are those with whom it has a defensive alliance pact whereby the partner state has made a promise to defend the state in question. Defensive alliances involve some form of structured defense cooperation through which states cooperate militarily or with defense capabilities beyond just mere coincidence. When a state formally agrees to defend another state in the event of attack, those states have a cooperative security relationship that suggests a mutual belief that the benefits of cooperation outweigh the risk of opportunism and the cost of coordination [@benson_assessingvariationformal_2016]. Data on state participation in defensive alliance pacts is provided by the Alliance Treaty and Provisions (ATOP) data set, version 5 [@leeds_alliancetreatyobligations_2002].

I include a set of control variables that existing theories indicate could be causally related to the dependent and/or independent variables of interest. The models control for regime type, coding a country-year as a democracy if they score higher than 6 on the 21-point Polity V index. Democracies may build more capital intensive militaries because of casualty sensitivity [@gartzke_democracypreparationwar_2001; @caverley_democraticmilitarismvoting_2014], spend less on defense [@fordham_kantianliberalismregime_2005], are generally less likely to be involved in conflicts [@russett_triangulatingpeacedemocracy_2000], and may be more or less reliable alliance partners [@gartzke_whydemocraciesmay_2004; @digiuseppe_armsdemocraticallies_2016]. There is also a control for whether a country has been involved in an interstate war in the previous half decade, as states currently or recently engaged in conflicts may have different military capabilities than those not facing a salient military threat [@ghosn_mid3dataset_2004] and recent conflict experience may change patterns of innovation [@kollars_warhorizonsoldierled_2015]. The model controls for GDP, as resource-constrained states may be forced to specialize and may also be unable to replace labor with capital [@diehl_substitutescomplementseffects_1994; @craig_measuringtechnologicalcapability_2018].[^9] Finally, I control for CINC scores, as some states may harbor more global ambitions for which power projection capabilities are conducive and states with smaller militaries have greater incentives to specialize since the economic benefits of doing so are more salient [@markowitz_powerproximitydemocracy_2018].

[^9]: Others question whether we empirically observed more specialization by states facing higher budget constraints. In the European context, many states have diversified their military portfolios despite the financial cost of doing so [@howorth_securitydefencepolicy_2007; @janning_greatwideopen_2007].

## Model and Results

The dependent variable is military specialization of country $i$ in year $j$, measured with the functional entropy index described above. Higher values indicate more specialization and less diversification. As the dependent variable is a continuous measure, I estimate a series of ordinary least square (OLS) regressions. I estimate a series of models using the two different independent variables - (1) logged sum of allied military spending, (2) ratio of a country's CINC score to that of all its allies and itself. For each independent variable, I estimate a bivariate model then a model with all control variables plus temporal scaled cubic polynomials and country-clustered standard errors. Year scaled cubic polynomials appropriately account for temporal trends in specialization; technology evolves over time and militaries innovate; we should not expect the make up a state's military in 1980 to be composed of the same assets as 1990 [@carter_backfuturemodeling_2010]. Country-clustered standard errors account for the non-independence between observations in panel data [@cameron_practitionerguideclusterrobust_2015].

@tbl-results shows the results of all four models. The results from Models 1 and 2 demonstrate that allied military spending is positively associated with military specialization with statistical significance of at least the 0.05 standardized level. Although allies' CINC ratio is negative associated with military specialization (Model 3), the inclusion of control variables and temporal dependencies (Model 4) reverses the association and provides results that are consistent with the other models and theoretical expectations. In aggregate, these results provides suggestive evidence that states that have militarily capable alliance partners have more specialized military portfolios - omitting certain capabilities and over-producing other capabilities - relative to states that are reliant upon self-defense. We present the OLS results here as they are most easily interpretable and consistent with modeling specifications in existing research with similar data [@brownlee_cognitiveshortcutspublic_2019]. Nonetheless, these results are robust to a series of alternate model specifications provided in the appendix that relax assumptions about the underlying distribution of the dependent variable and temporal and country-specific trends. Similar results are found using fractional logit and beta regressions, Bayesian zero-inflated and ordered beta regressions, and with alternate specifications for year-fixed effects and robust clustering of standard errors.

```{r}
#| label: tbl-results
#| tbl-cap: "Model results"
#| fig.position: 'H!'

# EV 1
## Select vars
df_m1 <- df_full %>%
  dplyr::select(spec_fentrop,
                milex_allies,
                democracy,
                interstatewar_5yrlag_binary,
                lngdp_WDI_full,
                cinc_MC,
                year_base,
                year_sq,
                year_cube)

## Bivariate
m1_bi <- lm(spec_fentrop ~ milex_allies,
         data = df_m1)
m1_bicse <- lmtest::coeftest(m1_bi, 
                              vcov = sandwich::vcovCL(m1_bi,
                                                      factor(df_full$country)))

## Multivariate
m1_spline <- lm(spec_fentrop ~ .,
                data = df_m1)
m1_splinecse <- lmtest::coeftest(m1_spline,
                                 vcov = sandwich::vcovCL(m1_spline, factor(df_full$country)))

# EV 2
## Select vars
df_m2 <- df_full %>%
  dplyr::select(spec_fentrop,
                cinc_allies_ratio,
                democracy,
                interstatewar_5yrlag_binary,
                lngdp_WDI_full,
                cinc_MC,
                year_base,
                year_sq,
                year_cube)

## Bivariate
m2_bi <- lm(spec_fentrop ~ cinc_allies_ratio,
            data = df_m2)
m2_bicse <- lmtest::coeftest(m2_bi, 
                              vcov = sandwich::vcovCL(m2_bi,
                                                      factor(df_full$country)))

## Multivariate
m2_spline <- lm(spec_fentrop ~ .,
                data = df_m2)
m2_splinecse <- lmtest::coeftest(m2_spline,
                                 vcov = sandwich::vcovCL(m2_spline, factor(df_full$country)))

# Table
## https://posit.co/blog/new-features-upgrades-in-gt-0-8-0/
models <- list(m1_bicse, m1_splinecse,
               m2_bicse, m2_splinecse)

coefs <- c("milex_allies" = "Allies' Mil Spend. (log)",
           "cinc_allies_ratio" = "Allies' CINC Ratio",
           "democracy1" = "Democracy",
           "interstatewar_5yrlag_binary1" = "Interstate War (5yr lag)",
           "lngdp_WDI_full" = "GDP (log)",
           "cinc_MC" = "CINC",
           "year_base" = "Year",
           "year_sq" = "Year$^2$",
           "year_cube" = "Year$^3$")

modelsummary::modelsummary(models,
                           statistic = 'p.value',
                           stars = TRUE,
                           coef_map = coefs,
                           escape = FALSE) %>%
  kableExtra::add_footnote(label = "All models include country-clustered standard errors.")
```

```{r subst-fx}
subst <- marginaleffects::comparisons(m2_spline,
                                      variables = list(cinc_allies_ratio = "sd")) %>%
  summary() %>%
  dplyr::pull(estimate) / max(df_m1$spec_fentrop, na.rm = TRUE) * 100
```

The relationship between alliances and military specialization is also substantively significant. Holding all control variables constant, a one standard deviation increase in allies' CINC ratio (independent variable in Models 3 and 4) is associated with a `r round(subst, 1)`% increase in a state's military specialization; which is how much more specialized Japan's military became between 1982 and 2000. Despite what what is traditionally understood as a lopsided division of security responsibilities, US and Japanese security responsibilities have been purposefully divided [@ando_empiricalanalysisdefense_2015]. Japan's 1982 capability realignment described in @sec-intro signaled the start of a new era of coordinated military cooperation with the United States, with the joint communique issued by Prime Minister Suzuki and President Reagan [-@usdepartmentofstate_visitjapaneseprime_1981, 3] stressing "the desirability of an appropriate division of roles between Japan and the United States". Japan was entrusted with protecting its sea lines of communication (SLOCs) 1,000 nautical miles off its coast and providing logistical support to offensive US operations as needed. @fig-japan illustrates how one result of the alliance relationship was a more specialized Japanese distribution of military capabilities. Japan doubled its SAM and far-from-shore naval capabilities like destroyers and utility helicopters and significantly downsized its amphibious and coastal fleets. The alliance relationship with the United States allowed Japan to carry the "defensive shield" by specializing in capabilities for SLOCs and "rear-area support" while forgoing attack-capable surface ships and high-tech long-range aircraft where the United States excelled by serving as the alliance's "offensive spear" [@schoff_howupgradejapan_2014].

```{r}
#| label: fig-japan
#| fig-cap: "Change in Japan's distribution of military capabilities between 1982 and 2000. Capabilities Japan did not possess at any point during this time period (eg ballistic missiles and drones) are omitted."
#| fig-height: 8
#| fig-width: 6
#| fig-pos: H

df_long %>%
  tidyr::separate(tek,
                  into = c("parent", "child"),
                  sep = "_",
                  remove = FALSE) %>%
  dplyr::group_by(year, tek) %>%
  dplyr::mutate(percentile = round(dplyr::percent_rank(unit_count), 2)) %>%
  dplyr::ungroup() %>%
  dplyr::filter(country == "japan" &
                  year %in% c(1982, 2000)) %>%
  dplyr::mutate(parent = dplyr::recode(parent,
                                       'principal surface combatants' = 'surface ships',
                                       'patrol and coastal combatants' = 'coastal ships',
                                       'anti-tank/anti-infrastructure' = 'anti-tank',
                                       'armoured fighting vehicles' = 'combat vehicle',
                                       'unmanned aerial vehicles' = 'uav'),
                child = dplyr::recode(child,
                                      'maritime patrol' = 'maritime',
                                      'mine countermeasures' = 'minesweepers',
                                      'paramilitary' = 'paramil.',
                                      'recon-surveil' = 'recon',
                                      'surveillance and intelligence' = 'surveillance',
                                      'surface to air missiles' = 'sam',
                                      'search and rescue' = 'sar',
                                      'surface to surface missiles' = 'missiles',
                                      'surface to surface artillery' = 'artillery')) %>%
  dplyr::mutate(parent = stringr::str_to_title(parent),
                child = stringr::str_to_title(child),
                child = dplyr::recode(child,
                                      'Aew' = 'AEW',
                                      'Ew' = 'EW',
                                      'Sam' = 'SAM',
                                      'Sar' = 'SAR')) %>%
  dplyr::filter(!parent %in% c("Ballistic Missiles",
                               "Land/Sea Defence",
                               "Logistics And Support",
                               "Radars",
                               "Uav")) %>%
  dplyr::filter(!tek %in% c("air defence_surface to air artillery",
                            "air defence_air to air missiles",
                            "aircraft_ew",
                            "aircraft_attack",
                            "aircraft_aew",
                            "aircraft_bomber",
                            "aircraft_maritime patrol",
                            "aircraft_paramilitary",
                            "aircraft_recon-surveil",
                            "amphibious_other",
                            "anti-tank/anti-infrastructure_missiles",
                            "armoured fighting vehicles_engineering and maintenance",
                            "helicopters_ew",
                            "helicopters_maritime",
                            "helicopters_observation",
                            "land/sea defence_air to surface ordnance",
                            "mine warfare_mine layers",
                            "patrol and coastal combatants_other",
                            "patrol and coastal combatants_coastal",
                            "submarines_other",
                            "submarines_coastal")) %>%
  dplyr::arrange(tek, year) %>%
  dplyr::mutate(paired = rep(1:(dplyr::n()/2), each = 2),
                year = factor(year)) %>%
  dplyr::group_by(paired) %>%
  dplyr::mutate(diffval = unit_count - dplyr::lag(unit_count), default = dplyr::first(unit_count)) %>%
  ggplot(aes(x = unit_count,
             y = child)) +
  geom_path(aes(group = paired),
            data = ~subset(., diffval > 0 |
                             diffval < 0 |
                             is.na(diffval)),
            arrow = arrow(angle = 20,
                          length = unit(0.1, "in"),
                          type = "open")) +
  geom_point(aes(color = year), size = 3) +
  facet_wrap(~ parent,
             ncol = 2,
             scales = "free") +
  scale_x_continuous(breaks = scales::pretty_breaks(),
                     labels = scales::label_number(big.mark = ",")) +
  theme_bw() +
  theme(text = element_text(size = 11),
        plot.title = element_markdown(),
        legend.position = "none",
        legend.title = element_blank(),
        axis.title.y = element_blank()) +
  scale_color_brewer(palette = "Paired") +
  labs(title = "Japan's Distribution of Military Capabilities in <span style = 'color: #A6CEE3;'><b>1982</b></span> and <span style = 'color: #1F78B4;'><b>2000</b></span>",
       x = "Count")
```

# Conclusion {#sec-conclusion}

"Make-or-buy" in the context of international security is better thought of as "make/buy-or-rely". States can "make/buy" and provide for their security on their own in the anarchic self-help international system or they can "rely" and pool resources with others. In relying on militarily capable allies, states may rationally opt out of diversifying their military. Even if states' fear of exploitation is most salient where survival may be at stake, military alliances can sufficiently manage uncertainty about cooperation under anarchy by providing gains to cooperation that exceed the expected cost of opportunism and coordination. In doing so states can specialize their own militaries, garnering specialization's efficiency benefits while minimizing its efficacy costs.

A shared production model of military capabilities does not necessarily run counter to conventional wisdom advocating military portfolio diversification; it simply represents a re-framing of the unit of analysis. Allies influence armaments. A state's security can be improved not just by their own defense efforts, but those of similarly aligned states [@yarhi-milo_armallypatron_2016]. Without discounting the role that domestic and external threat considerations play in armament decisions, this paper's contribution is explaining how states differ in the composition of their military portfolios and the role that alliances play in explaining some of that variation.

Whether it is wise for states to omit or overproduce military capabilities on the basis of their alliance relationships is a separate question this paper has not tried to answer. Specialization may make states better off, since they can "take advantage of economies of scale in the provision of defense and to benefit from specialization by coordinating training, equipment, and procedures. By pooling their efforts and/or cooperating with states that have different comparative advantages, leaders hope to create a stronger joint fighting force [@leeds_allianceinstitutionalizationalliance_2005, 185]. On the other hand,"fear that the other will not live up to the terms of the agreement" [@lake_anarchyhierarchyvariety_1996, 15] may be well-warranted and, importantly, consequential for how well a state fares when it actually comes time to partner together on the battlefield [@cranmer_coalitionqualitymultinational_2018].

A few shortcomings in this paper provide avenues for future research. There are a variety of forms of governance that allows states to "rely" to different degrees and for different reason; the analysis here simply looks at defense pacts and operationalizes reliability in alignment with existing research. But differences in interest alignment, symmetry, and coercion across alliances may influence who specializes in what [@lanoszka_consentcoercionusing_2013]. Future work could investigate the characteristics of an alliance that are most conducive to this division of labor by looking at these factors as well as different types of security alignments like defense cooperation agreements (DCAs) [@kinne_defensecooperationagreement_2020], how alliance treaties are negotiated [@poast_arguingalliancesart_2019, 174-175], long-term alliances versus temporary coalitions [@kreps_coalitionsconvenienceunited_2011], and the number of alliance partners [@fordham_allalliancesare_2016].

The insights here can help inform current debates about changing alliance relations and identify the consequences of allies trusting each other less than they used to. These debates often turn to the question about whether allies are contributing enough to the alliance. But by looking at *what* states are contributing to the common defense, rather than *how much* they are spending, new perspectives on burden sharing and the value of the alliance emerge [@gannon_whatcanfinland_2022]. Japan's defense strategy has not remained static, and contemporary discussions about their increased militarization in response to Chinese, North Korean, and Russian threat have put the US-Japan alliance front and center [@cooper_japanshiftwar_2023]. On the European front, then US ambassador to NATO Ivo @daalder_renewedambitionsnato_2013 gave a speech in which he noted that the problem was not that NATO countries were not spending enough money on defense, it was that they were not spending that money wisely. In both cases, discussions surrounding the US relationship with its allies has largely still concerned burden-sharing *costs* rather than burden-sharing *configurations*. An ally's defense spending matters because of what military capabilities they spend those resources on. If US allies feel confident they can rely on the United States, we should see them continue to specialize their militaries. Conversely, if US allies begin diversifying their military portfolios, that may be evidence that they are hedging their bets in seeking to defend themselves with a diversified full-spectrum force rather than counting on the United States to provide capabilities they are omitting.

\newpage

# References
